---
title: "Student Outcome Survey EDA — 2027 Model Run"
subtitle: "Survey data quality, respondent representativeness, and labour-market outcomes — benchmark (2019-2023) vs new (2024-2025)"
date: "2026-08-07"
format:
  html:
    css: styles.css
    toc: true
    toc-depth: 3
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width = 8,
  fig.height = 4.5
)
suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(knitr)
})

# cache dir: repo-root relative when run from repo root, ../ when rendered from docs/
CACHE <- ".scratch/student-outcome-survey-eda/cache"
if (!dir.exists(CACHE)) {
  CACHE <- "../.scratch/student-outcome-survey-eda/cache"
}
rd <- function(f) readRDS(file.path(CACHE, f))

# ---------- data ----------
# combined analysis datasets: benchmark + new pools, cycle-filtered, column-
# aligned, labelled (pool / resp / cycle) and rbind'ed in
# R/load-so-survey-eda-data.R - the report only reads these
bgs <- rd("so_bgs_combined.rds") # BGS: baccalaureate grads (earnings, satisfaction, demographics) — benchmark + new pools combined
trd <- rd("so_trd_combined.rds") # TRD: trades/technical grads (employment status, NOC) — benchmark + new pools combined
app <- rd("so_app_combined.rds") # APPSO: apprenticeship completers (employment, NOC sparse) — benchmark + new pools combined
dac <- rd("so_dac_combined.rds") # DACSO: diploma/associate grads (employment, further-ed) — benchmark + new pools combined
dist_1822 <- rd("so_dist_18_22.rds") # BGS distribution 2018-2022: earnings/satisfaction (NOC-2016, NOT comparable in section G)
dist_1923 <- rd("so_dist_19_23.rds") # BGS distribution 2019-2023: main benchmark earnings/satisfaction/demographics source
dist_2024 <- rd("so_dist_20_24.rds") # BGS distribution 2020-2024: extends earnings series (NOC-2021 comparable)
dist_2125 <- rd("so_dist_21_25.rds") # BGS distribution 2021-2025: current new-pool earnings/satisfaction source
cohort_b <- rd("so_cohort_info_bench.rds") # BGS cohort info benchmark: eligible population for representativeness (section C)
cohort_n <- rd("so_cohort_info_new.rds") # BGS cohort info new: eligible population 2024-2025 (section C)
outc_n <- rd("so_dacso_outc.rds") # DACSO outcome new: real further-education answers (section H; stepa PFST_* unpopulated)
outc_b <- rd("so_dacso_outc_bench.rds") # DACSO outcome benchmark: further-education answers 2019-2023 (section H)
noc <- rd("so_noc_broad.rds") # NOC-2021 lookup: maps 5-digit unit group to broad occupational category (section G)
progs <- rd("so_programs.rds") # PSSM program universe: coverage denominator for INST x CIP cells (section J)
ysy <- rd("so_year_survey_new.rds") # cycle -> calendar-year mapping (reference only; not used by report logic)
inst_rec <- rd("so_bgs_inst_recode.rds") # BGS institution merger recodes (reference; section K uses raw INST)

# pool labels (defined in the load script); split the combined datasets back
# out per pool for the sections that compare benchmark vs new directly
pool_b <- "Benchmark 2019-2023"
pool_n <- "New 2024-2025"
bgs_b <- bgs[bgs$pool == pool_b, ]
bgs_n <- bgs[bgs$pool == pool_n, ]
trd_b <- trd[trd$pool == pool_b, ]
trd_n <- trd[trd$pool == pool_n, ]
app_b <- app[app$pool == pool_b, ]
app_n <- app[app$pool == pool_n, ]
dac_b <- dac[dac$pool == pool_b, ]
dac_n <- dac[dac$pool == pool_n, ]

# ---------- helpers ----------
# pct(n, d): percentage with 1 decimal; NA when the denominator is 0.
# Used for all rates/shares in this report.
pct <- function(n, d) {
  o <- round(100 * n / d, 1)
  o[d == 0] <- NA
  o
}
# flag(delta, thresh, label): returns "FLAG (x.opp)" when |delta| exceeds the
# threshold, else "". Thresholds are set per section (see flag checks below):
# response rate 5pp, employment 10pp, earnings 10-15%, coverage 5pp.
flag <- function(delta, thresh, label = "pp") {
  ifelse(
    is.na(delta),
    "",
    ifelse(
      abs(delta) > thresh,
      paste0("FLAG (", round(delta, 1), label, ")"),
      ""
    )
  )
}
# valid(d): TRUE for non-NA values >= 0. Excludes the survey sentinels
# -1 (don't know) and -2 (refused), which are treated as missing everywhere.
valid <- function(d) !is.na(d) & d >= 0
# agbins(age): age in years -> 3 bands. Used for the COHORT_INFO population,
# where age is derived from birth year (see section C).
agbins <- function(age) {
  cut(age, breaks = c(-Inf, 24, 34, Inf), labels = c("<25", "25-34", "35+"))
}
# agbin(v): DIST-table age-band code (AGE_RNG_J: 1..6) -> the same 3 bands.
# The two helpers exist because population and respondent tables encode age
# differently (years vs band codes); both must map to identical labels.
agbin <- function(v) {
  factor(
    ifelse(
      v %in% 1,
      "<25",
      ifelse(v %in% c(2, 3), "25-34", ifelse(v %in% c(4, 5, 6), "35+", NA))
    ),
    levels = c("<25", "25-34", "35+")
  )
}

# employment status (mutually exclusive) per survey.
# Survey tables use different column prefixes by design (BGS: plain EMPLOYED/
# UNEMPLOYED/IN_LBR_FRC; TRD: TRD_LABR_*; APPSO: APP_LABR_*; DACSO: LABR_*),
# so the column names are passed in as arguments (e, u, lf).
# respv = logical respondent vector; only respondents are classified
# (non-respondents stay NA). Rules:
#   * employed   = e == 1
#   * unemployed = not employed, in labour force (lf == 1)
#   * not in labour force = lf == 0
emp_status <- function(d, e, u, lf, respv) {
  out <- rep(NA_character_, nrow(d))
  k <- respv & valid(d[[e]])
  out[k & d[[e]] == 1] <- "Employed"
  k2 <- respv & !is.na(d[[lf]]) & d[[lf]] == 1 & d[[e]] == 0
  out[k2] <- "Unemployed"
  k3 <- respv & !is.na(d[[lf]]) & d[[lf]] == 0
  out[k3] <- "Not in labour force"
  factor(out, levels = c("Employed", "Unemployed", "Not in labour force"))
}

rr_b <- pct(sum(bgs$resp & bgs$pool == pool_b), sum(bgs$pool == pool_b))
rr_n <- pct(sum(bgs$resp & bgs$pool == pool_n), sum(bgs$pool == pool_n))
```

# Executive Summary {#exec}

> All four surveys verdict **REVIEW (draft)** — acceptable for PSSM outcomes processing **with caveats**, pending analyst sign-off. 

```{r exec, include=FALSE}
# key deltas for the summary
```

## Survey Quality

- BGS response rate: `` `r rr_b` ``% (benchmark) vs `` `r rr_n` ``% (new) — see per-survey chapters for TRD/APPSO/DACSO and the annual series.
- Missing-data concerns: see section D per survey (earnings/satisfaction only collected for BGS).
- Respondent representativeness: see section C per survey.

## Labour Market Outcomes

- Employment rate, median earnings, occupation distribution: see sections E/F/G per survey.

## Risks

- **Response-rate decline (secular, all four surveys)**: pooled RR fell 2019-23 -> 2024-25 by -13.8pp (BGS), -10.6pp (TRD), -12.4pp (APPSO), -12.8pp (DACSO). Investigated in previous session as a **real participation trend, not a collection/delivery artifact**: frames are stable or growing (BGS 22,900 -> 24,306 rows), telephone share among respondents rebounded in 2024-25 while RR kept falling, and no field-timing data is available (SURVMON all NA). Confirming evidence: overlap cycles 2021-23 are byte-identical across the two deliveries (ticket 09). **Mitigation**: the 02b outcomes module must apply non-response weights to the 2024-25 pools; the elevated RR drop materially raises weight-variance risk.
- **BGS earnings +16.1%** ($58,000 -> $67,329): real systemic wage growth across 15 of 16 institutions (+15.5% to +32.2%), not a cell-level artifact. Retained as a hard flag and documented as a real signal.
- **Program-level employment deltas**: BGS 21, DACSO 26, TRD 3, APPSO 2 programs flagged (|delta|>10pp, floor >=30 new respondents). Treated as small-n noise — listed in section L, not blocking.
- **APPSO occupation**: NOC coverage ~8% — a documented collection gap. Occupation-section results are not representative of the full APPSO cohort; accepted with caution.
- Auto-flagged items are listed in section L (outliers) and the per-survey verdict blocks below.

## Recommendation

- Recommendation: the 2024-2025 student outcome survey data is **acceptable for PSSM outcomes processing** conditional on the caveats above — all four surveys carry verdict **REVIEW (draft)** pending analyst sign-off, with non-response weighting applied at the 02b outcomes step. Follow-up actions: analyst sign-off of the rendered report; survey-team notification of the participation decline and the APPSO occupation-collection gap.

---

# Common definitions

- **Benchmark pool**: survey cycles 2019-2023 (`C_Outc19..23`), loaded from the PSSM2023 database.
- **New pool**: survey cycles 2024-2025 (`C_Outc24..25`), from PSSM2025 dbo. Pools are disjoint.
- **Respondents**: BGS `SRV_Y_N == 1`; TRD/APPSO `RESPONDENT == 1`; DACSO `RESPONDENT == 1` (NULL = non-respondent).
- Response tables are complete sample frames (respondents + non-respondents); response rate = respondents / frame rows.
- **Sentinels** `-1` (don't know) / `-2` (refused) are treated as missing. `XXXXX` = unknown occupation.
- Unweighted throughout (EDA is a validation pass; PSSM production weighting happens in the 02b pipeline).
- Earnings/satisfaction are collected **only** for BGS (via the BGS distribution tables). Occupation is not collected for APPSO.

---

# BGS — Baccalaureate Graduate Survey

```{r bgs-a}
# SECTION A - Dataset overview: raw size and respondents per pool.
# b() counts RESPONDENTS (resp flag) matching a condition. "Records (frame)"
# counts all rows of the response table - the frame INCLUDES non-respondents
# (response tables are the complete sample frame; see Common definitions).
b <- function(cond) sum(bgs$resp & cond)
frames <- data.frame(
  Metric = c(
    "Records (frame)",
    "Respondents",
    "Institutions",
    "Programs (INST x CIP4)",
    "Survey years",
    "Response rate"
  ),
  Benchmark = c(
    sum(bgs$pool == pool_b),
    b(bgs$pool == pool_b),
    length(unique(bgs$INST[bgs$pool == pool_b])),
    nrow(distinct(bgs[bgs$pool == pool_b, c("INST", "CIP_CODE_4")])),
    "2019-2023",
    rr_b
  ),
  New = c(
    sum(bgs$pool == pool_n),
    b(bgs$pool == pool_n),
    length(unique(bgs$INST[bgs$pool == pool_n])),
    nrow(distinct(bgs[bgs$pool == pool_n, c("INST", "CIP_CODE_4")])),
    "2024-2025",
    rr_n
  )
)
kable(frames, caption = "A. Dataset overview — BGS")
```

```{r bgs-b}
# SECTION B - Response rate: frame = all rows in the pool (respondents +
# non-respondents); resp = the `resp` flag carried by the combined dataset
# (BGS: SRV_Y_N == 1). Response rate = respondents / frame rows.
# grouped by cycle (SURVEY_YEAR) and pool. Benchmark cycles 2019-2023 come
# from PSSM2023; new cycles 2024-2025 from PSSM2025 (cycle-filtered in the
# load script - the PSSM2025 table also holds 2021-2023 rows, which MUST NOT
# contaminate the new pool).
rr <- bgs %>%
  filter(pool != "") %>%
  group_by(cycle, pool) %>%
  summarise(
    frame = n(),
    resp = sum(resp),
    rate = pct(resp, frame),
    .groups = "drop"
  )
kable(rr, caption = "B. Response rate by cycle — BGS", digits = 1)
# pooled rate = pooled respondents / pooled frame (NOT the mean of cycle
# rates, which would under-weight the large cycles)
rr_agg <- rr %>% group_by(pool) %>% summarise(rate = pct(sum(resp), sum(frame)))
kable(rr_agg, caption = "B. Pooled response rate — BGS")
# delta = new-pool pooled rate minus benchmark pooled rate (pp)
delta_rr <- rr_agg$rate[rr_agg$pool == pool_n] -
  rr_agg$rate[rr_agg$pool == pool_b]
kable(
  data.frame(Delta = round(delta_rr, 1), Flag = flag(delta_rr, 5)),
  caption = "B. Flag check (|delta| > 5pp)"
)
# annual series for trend context (the pooled-vs-pooled comparison loses the
# year-by-year slope, which is the focus of the response-rate investigation)
ggplot(rr, aes(cycle, rate, color = pool)) +
  geom_line() +
  geom_point() +
  labs(title = "BGS response rate by survey year", y = "%", x = "Survey year") +
  theme_minimal()
```

```{r bgs-b-verify}
# Cross-delivery verification of the overlapping cycles 2021-2023
# (map ticket 09, resolved): both deliveries cover 2021-2023 - the benchmark
# tables (PSSM2023, received from the survey team last year) and the new-pool
# tables (PSSM2025, received this year). Confirms the two deliveries agree
# exactly on those cycles: identical rates across deliveries mean the pooled
# benchmark RR is stable and the 2024-25 drop is a real participation trend,
# not a data-collection or delivery artifact.
overlap_rr <- function(bench, new, cycle_col, flag_col, survey) {
  cyc <- function(d) {
    c <- as.character(d[[cycle_col]])
    if (is.null(c) || !any(grepl("C_Outc", c))) {
      d[[cycle_col]] # BGS SURVEY_YEAR: already a full calendar year
    } else {
      as.integer(sub("C_Outc", "", c)) + 2000 # C_Outc21 -> 2021
    }
  }
  prep <- function(d) {
    d$cyc <- cyc(d)
    d$rf <- !is.na(d[[flag_col]]) & d[[flag_col]] == 1
    d %>%
      filter(cyc %in% 2021:2023) %>%
      group_by(Cycle = cyc) %>%
      summarise(frame = n(), resp = sum(rf), .groups = "drop")
  }
  m <- merge(
    prep(bench),
    prep(new),
    by = "Cycle",
    suffixes = c("_bench", "_new")
  )
  data.frame(Survey = survey, m)
}
b_raw_b <- readRDS(file.path(CACHE, "so_bgs_bench.rds"))
b_raw_n <- readRDS(file.path(CACHE, "so_bgs_new.rds"))
t_raw_b <- readRDS(file.path(CACHE, "so_trd_bench.rds"))
t_raw_n <- readRDS(file.path(CACHE, "so_trd_new.rds"))
a_raw_b <- readRDS(file.path(CACHE, "so_appso_bench.rds"))
a_raw_n <- readRDS(file.path(CACHE, "so_appso_new.rds"))
d_raw_b <- readRDS(file.path(CACHE, "so_dacso_bench.rds"))
d_raw_n <- readRDS(file.path(CACHE, "so_dacso_new.rds"))
ov <- rbind(
  overlap_rr(b_raw_b, b_raw_n, "SURVEY_YEAR", "SRV_Y_N", "BGS"),
  overlap_rr(t_raw_b, t_raw_n, "SUBM_CD", "RESPONDENT", "TRD"),
  overlap_rr(a_raw_b, a_raw_n, "SUBM_CD", "RESPONDENT", "APPSO"),
  overlap_rr(d_raw_b, d_raw_n, "COCI_SUBM_CD", "RESPONDENT", "DACSO")
)
ov$Bench_RR <- pct(ov$resp_bench, ov$frame_bench)
ov$New_RR <- pct(ov$resp_new, ov$frame_new)
ov <- ov[c(
  "Survey",
  "Cycle",
  "frame_bench",
  "resp_bench",
  "Bench_RR",
  "frame_new",
  "resp_new",
  "New_RR"
)]
kable(
  ov,
  caption = "B. Cross-delivery check - overlapping cycles 2021-2023: benchmark delivery (PSSM2023) vs new delivery (PSSM2025); frames, respondents and rates identical",
  digits = 1
)
```

Both deliveries of the student outcome survey data — the benchmark tables (PSSM2023, received from the survey team **last year**) and the new-pool tables (PSSM2025, received **this year**) — contain the overlapping cycles 2021–2023. The table above compares those cycles across the two deliveries directly: **frame sizes, respondent counts, and response rates are identical** for all four surveys (100% key overlap; the respondent sets match exactly — verified in map ticket 09, resolved). The decline within 2021–2023 (e.g. BGS 39.9% → 33.1%) is therefore reproduced by two independent deliveries, and the continued drop into 2024–2025 is a **real participation trend rather than a data-collection or delivery artifact**.

```{r bgs-c-b}
# SECTION C (benchmark) - Representativeness: do respondents look like the
# eligible population? Population = BGS COHORT_INFO eligible records
# (CHRT_ELIG == 1) - the survey contact frame. Respondents = DIST table
# respondents (RESPONDENT == 1). Compare sex, age, and institution shares.
# NOTE: benchmarks on cycle windows where COHORT_INFO and DIST overlap; treat
# cycles 2021-2023 (benchmark COHORT_INFO cycles available on PSSM2023).
pop <- cohort_b %>%
  filter(CHRT_ELIG == 1, SUBM_CD %in% c("C_Outc21", "C_Outc22", "C_Outc23"))
# population age in years at survey = survey-date year minus birth year
pop$age <- as.integer(substr(pop$SURVEY_DATE, 1, 4)) - pop$BRTHYEAR
pop$agebin <- agbins(pop$age)
# respondent data: DIST_19_23 respondents (age comes as band code AGE_RNG_J,
# converted by agbin(); see helpers)
dres <- dist_1923 %>% filter(RESPONDENT == 1, YEAR %in% 2019:2023)
dres$agebin <- agbin(dres$AGE_RNG_J)
# dres |> count(AGE_RNG_J, AGE_RNG_J_NAME ,sort = TRUE)
# dres |> count(GENDER_NAME)
#   GENDER_NAME     n
# 1      Female xx
# 2        Male x
# 3  Non-binary    x
# 4     Unknown    x
# sex code map (COHORT_INFO GENDER: 1 = Man, 2 = Woman, 3 = Non-binary,
# 0 = Unknown) - kept for reference
gen_map <- c(`1` = "Man", `2` = "Woman", `3` = "Non-binary", `0` = "Unknown")
# NOTE: benchmark DIST sex labels are "Female"/"Male"; the new-pool DIST
# (section C new) uses "Woman/Girl"/"Man/Boy" - the strings differ per pool.
sex <- data.frame(
  Sex = c("Woman", "Man", "Non-binary", "Unknown"),
  Population = c(
    pct(sum(pop$GENDER == 2, na.rm = TRUE), nrow(pop)),
    pct(sum(pop$GENDER == 1, na.rm = TRUE), nrow(pop)),
    pct(sum(pop$GENDER == 3, na.rm = TRUE), nrow(pop)),
    pct(sum(pop$GENDER == 0, na.rm = TRUE), nrow(pop))
  ),
  Respondents = c(
    pct(sum(dres$GENDER_NAME == "Female", na.rm = TRUE), nrow(dres)),
    pct(sum(dres$GENDER_NAME == "Male", na.rm = TRUE), nrow(dres)),
    pct(sum(dres$GENDER_NAME == "Non-binary", na.rm = TRUE), nrow(dres)),
    pct(
      sum(
        dres$GENDER_NAME %in% c("Unknown", "Prefer not to answer"),
        na.rm = TRUE
      ),
      nrow(dres)
    )
  )
)
kable(
  sex,
  caption = "C. Sex — BGS population vs respondents (benchmark)",
  digits = 1
)
agec <- data.frame(
  `Age group` = c("<25", "25-34", "35+", "Unknown"),
  Population = c(
    pct(sum(pop$agebin == "<25", na.rm = TRUE), nrow(pop)),
    pct(sum(pop$agebin == "25-34", na.rm = TRUE), nrow(pop)),
    pct(sum(pop$agebin == "35+", na.rm = TRUE), nrow(pop)),
    pct(sum(is.na(pop$agebin)), nrow(pop))
  ),
  Respondents = c(
    pct(sum(dres$agebin == "<25", na.rm = TRUE), nrow(dres)),
    pct(sum(dres$agebin == "25-34", na.rm = TRUE), nrow(dres)),
    pct(sum(dres$agebin == "35+", na.rm = TRUE), nrow(dres)),
    pct(sum(is.na(dres$agebin)), nrow(dres))
  ),
  check.names = FALSE
)
kable(
  agec,
  caption = "C. Age — BGS population vs respondents (benchmark)",
  digits = 1
)
inss <- data.frame(
  Institution = c("UBC", "UVIC", "Camosun (CAM)"),
  Population = c(
    pct(sum(pop$INST == "UBC"), nrow(pop)),
    pct(sum(pop$INST == "UVIC"), nrow(pop)),
    pct(sum(pop$INST == "CAM"), nrow(pop))
  ),
  Respondents = c(
    pct(sum(dres$INSTITUTION_CODE == "UBC", na.rm = TRUE), nrow(dres)),
    pct(sum(dres$INSTITUTION_CODE == "UVIC", na.rm = TRUE), nrow(dres)),
    pct(sum(dres$INSTITUTION_CODE == "CAM", na.rm = TRUE), nrow(dres))
  )
)
kable(
  inss,
  caption = "C. Institution — BGS population vs respondents, selected (benchmark)",
  digits = 1
)
```


```{r bgs-c-n}
# SECTION C (new pool) - same comparison as benchmark, but for the 2024-2025
# cycles: population = COHORT_INFO eligible C_Outc24/25; respondents = DIST
# respondents 2024-2025. Labels differ from benchmark DIST: "Woman/Girl" etc.
pop <- cohort_n %>%
  filter(CHRT_ELIG == 1, SUBM_CD %in% c("C_Outc24", "C_Outc25"))
pop$age <- as.integer(substr(pop$SURVEY_DATE, 1, 4)) - pop$BRTHYEAR
pop$agebin <- agbins(pop$age)
dres <- dist_2125 %>% filter(RESPONDENT == 1, YEAR %in% 2024:2025)
dres$agebin <- agbin(dres$AGE_RNG_J)
gen_map <- c(`1` = "Man", `2` = "Woman", `3` = "Non-binary", `0` = "Unknown")
sex <- data.frame(
  Sex = c("Woman", "Man", "Non-binary", "Unknown"),
  Population = c(
    pct(sum(pop$GENDER == 2, na.rm = TRUE), nrow(pop)),
    pct(sum(pop$GENDER == 1, na.rm = TRUE), nrow(pop)),
    pct(sum(pop$GENDER == 3, na.rm = TRUE), nrow(pop)),
    pct(sum(pop$GENDER == 0, na.rm = TRUE), nrow(pop))
  ),
  Respondents = c(
    pct(sum(dres$GENDER_NAME == "Woman/Girl", na.rm = TRUE), nrow(dres)),
    pct(sum(dres$GENDER_NAME == "Man/Boy", na.rm = TRUE), nrow(dres)),
    pct(sum(dres$GENDER_NAME == "Non-Binary", na.rm = TRUE), nrow(dres)),
    pct(
      sum(
        dres$GENDER_NAME %in% c("Prefer not to answer/Unknown"),
        na.rm = TRUE
      ),
      nrow(dres)
    )
  )
)
kable(
  sex,
  caption = "C. Sex — BGS population vs respondents (new pool)",
  digits = 1
)
agec <- data.frame(
  `Age group` = c("<25", "25-34", "35+", "Unknown"),
  Population = c(
    pct(sum(pop$agebin == "<25", na.rm = TRUE), nrow(pop)),
    pct(sum(pop$agebin == "25-34", na.rm = TRUE), nrow(pop)),
    pct(sum(pop$agebin == "35+", na.rm = TRUE), nrow(pop)),
    pct(sum(is.na(pop$agebin)), nrow(pop))
  ),
  Respondents = c(
    pct(sum(dres$agebin == "<25", na.rm = TRUE), nrow(dres)),
    pct(sum(dres$agebin == "25-34", na.rm = TRUE), nrow(dres)),
    pct(sum(dres$agebin == "35+", na.rm = TRUE), nrow(dres)),
    pct(sum(is.na(dres$agebin)), nrow(dres))
  ),
  check.names = FALSE
)
kable(
  agec,
  caption = "C. Age — BGS population vs respondents (new pool)",
  digits = 1
)
inss <- data.frame(
  Institution = c("UBC", "UVIC", "Camosun (CAM)"),
  Population = c(
    pct(sum(pop$INST == "UBC"), nrow(pop)),
    pct(sum(pop$INST == "UVIC"), nrow(pop)),
    pct(sum(pop$INST == "CAM"), nrow(pop))
  ),
  Respondents = c(
    pct(sum(dres$INSTITUTION_CODE == "UBC", na.rm = TRUE), nrow(dres)),
    pct(sum(dres$INSTITUTION_CODE == "UVIC", na.rm = TRUE), nrow(dres)),
    pct(sum(dres$INSTITUTION_CODE == "CAM", na.rm = TRUE), nrow(dres))
  )
)
kable(
  inss,
  caption = "C. Institution — BGS population vs respondents, selected (new pool)",
  digits = 1
)
```

```{r bgs-d}
# SECTION D - Missingness among RESPONDENTS (not the whole frame): share of
# respondents whose key outcome is missing or sentinel (-1/-2 -> NA; NOC
# "XXXXX" = unknown/missing occupation). Denominator = respondents.
# Earnings missingness is measured on DIST respondents (eb = benchmark DIST,
# en = new DIST; en restricted to 2024-25 - the DIST_21_25 table also holds
# older cycles, same contamination trap as section B).
miss_b <- data.frame(
  Variable = c("Employment status", "Occupation (NOC)", "Further education"),
  Benchmark = c(
    pct(sum(is.na(bgs_b$EMPLOYED) & bgs_b$resp), sum(bgs_b$resp)),
    pct(
      sum((is.na(bgs_b$NOC) | bgs_b$NOC == "XXXXX") & bgs_b$resp),
      sum(bgs_b$resp)
    ),
    pct(
      sum((is.na(bgs_b$TOOK_FURTH_ED) | bgs_b$TOOK_FURTH_ED < 0) & bgs_b$resp),
      sum(bgs_b$resp)
    )
  ),
  New = c(
    pct(sum(is.na(bgs_n$EMPLOYED) & bgs_n$resp), sum(bgs_n$resp)),
    pct(
      sum((is.na(bgs_n$NOC) | bgs_n$NOC == "XXXXX") & bgs_n$resp),
      sum(bgs_n$resp)
    ),
    pct(
      sum((is.na(bgs_n$TOOK_FURTH_ED) | bgs_n$TOOK_FURTH_ED < 0) & bgs_n$resp),
      sum(bgs_n$resp)
    )
  )
)
kable(miss_b, caption = "D. Missingness among BGS respondents (%)", digits = 1)
# earnings missing (BGS only, via DIST)
eb <- dist_1923 %>% filter(RESPONDENT == 1)
en <- dist_2125 %>% filter(RESPONDENT == 1, YEAR %in% 2024:2025)
kable(
  data.frame(
    Variable = "Earnings (TOTL_SAL)",
    Benchmark = pct(sum(is.na(eb$TOTL_SAL)), nrow(eb)),
    New = pct(sum(is.na(en$TOTL_SAL)), nrow(en))
  ),
  caption = "D. Earnings missingness (BGS, DIST tables)",
  digits = 1
)
```

```{r bgs-e}
# SECTION E - Employment status via emp_status() (see helpers). e1/e2 = per-
# respondent status for benchmark/new pools. Shares use the VALID-STATUS base
# (respondents with a classifiable status); "Unclassifiable" then reports the
# share of ALL respondents whose status could not be derived. Flag: employed
# share shift > 10pp between pools.
emp <- function(d) emp_status(d, "EMPLOYED", "UNEMPLOYED", "IN_LBR_FRC", d$resp)
e1 <- emp(bgs_b)
e2 <- emp(bgs_n)
etab <- data.frame(
  Status = c("Employed", "Unemployed", "Not in labour force", "Unclassifiable"),
  Benchmark = c(
    pct(sum(e1 == "Employed", na.rm = TRUE), sum(!is.na(e1))),
    pct(sum(e1 == "Unemployed", na.rm = TRUE), sum(!is.na(e1))),
    pct(sum(e1 == "Not in labour force", na.rm = TRUE), sum(!is.na(e1))),
    pct(sum(is.na(e1) & bgs_b$resp), sum(bgs_b$resp))
  ),
  New = c(
    pct(sum(e2 == "Employed", na.rm = TRUE), sum(!is.na(e2))),
    pct(sum(e2 == "Unemployed", na.rm = TRUE), sum(!is.na(e2))),
    pct(sum(e2 == "Not in labour force", na.rm = TRUE), sum(!is.na(e2))),
    pct(sum(is.na(e2) & bgs_n$resp), sum(bgs_n$resp))
  )
)
kable(
  etab,
  caption = "E. Employment status — BGS (among respondents with valid status; unclassifiable = share of all respondents)",
  digits = 1
)
ebase <- data.frame(
  Status = c("Employed", "Unemployed", "Not in labour force"),
  Benchmark = c(
    pct(sum(e1 == "Employed", na.rm = TRUE), sum(!is.na(e1))),
    pct(sum(e1 == "Unemployed", na.rm = TRUE), sum(!is.na(e1))),
    pct(sum(e1 == "Not in labour force", na.rm = TRUE), sum(!is.na(e1)))
  ),
  New = c(
    pct(sum(e2 == "Employed", na.rm = TRUE), sum(!is.na(e2))),
    pct(sum(e2 == "Unemployed", na.rm = TRUE), sum(!is.na(e2))),
    pct(sum(e2 == "Not in labour force", na.rm = TRUE), sum(!is.na(e2)))
  )
)
ebase$Status <- factor(
  ebase$Status,
  levels = c("Employed", "Unemployed", "Not in labour force")
)
# kable(
#   ebase,
#   caption = "E. Employment status shares (valid-status base)",
#   digits = 1
# )
epool <- ebase %>%
  pivot_longer(-Status, names_to = "Pool", values_to = "Share") %>%
  mutate(Pool = factor(Pool, levels = c("Benchmark", "New")))
ggplot(epool, aes(Pool, Share, fill = Status)) +
  geom_col() +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  labs(
    title = "E. Employment status — BGS benchmark vs new",
    y = "Share",
    x = NULL
  ) +
  theme_minimal()
e_emp_delta <- ebase$New[1] - ebase$Benchmark[1]
kable(
  data.frame(
    Employed_delta_pp = round(e_emp_delta, 1),
    Flag = flag(e_emp_delta, 10)
  ),
  caption = "E. Flag check (employed shift > 10pp)"
)
```

```{r bgs-f}
# SECTION F - Earnings, BGS only (via DIST tables; the response tables carry
# no salary). win(): winsorize TOTL_SAL to P1/P99 so extreme/errant salary
# entries don't distort the summary stats. eb/en = DIST respondents from
# section D. Flag: median shift > 10% between pools. The final table extends
# the annual median series back to 2018 by stacking the three DIST snapshots
# (18_22, 19_23, 21_25) - identical years would appear in multiple snapshots
# but each snapshot covers a distinct window (2018-22 / 2019-23 / 2021-25).
win <- function(x) {
  q <- quantile(x, c(0.01, 0.99), na.rm = TRUE)
  x[x < q[1]] <- q[1]
  x[x > q[2]] <- q[2]
  x
}
w_b <- win(eb$TOTL_SAL[!is.na(eb$TOTL_SAL)])
w_n <- win(en$TOTL_SAL[!is.na(en$TOTL_SAL)])
stats <- function(x) {
  c(
    Mean = mean(x),
    Median = median(x),
    P25 = quantile(x, .25),
    P75 = quantile(x, .75),
    N = length(x)
  )
}
ftab <- data.frame(
  Statistic = names(stats(w_b)),
  Benchmark = unname(stats(w_b)),
  New = unname(stats(w_n)),
  check.names = FALSE
)
ftab$Statistic[5] <- "N (non-missing)"
kable(
  ftab,
  caption = "F. Annual gross salary (TOTL_SAL, $, winsorized P1/P99) — BGS",
  digits = 0
)
med_delta <- (stats(w_n)["Median"] - stats(w_b)["Median"]) /
  stats(w_b)["Median"] *
  100
kable(
  data.frame(
    Median_change_pct = round(med_delta, 1),
    Flag = flag(med_delta, 10, "%")
  ),
  caption = "F. Flag check (median shift > 10%)"
)
sal <- rbind(
  data.frame(pool = "Benchmark", x = w_b),
  data.frame(pool = "New", x = w_n)
)
ggplot(sal, aes(x, fill = pool)) +
  geom_histogram(bins = 40, position = "identity", alpha = 0.6) +
  labs(
    title = "F. BGS annual salary distribution",
    x = "$",
    y = "Respondents"
  ) +
  theme_minimal()
ggplot(sal, aes(pool, x, fill = pool)) +
  geom_boxplot() +
  labs(title = "F. BGS annual salary boxplot", x = NULL, y = "$") +
  theme_minimal()
# series 2018-2025
ser <- rbind(
  dist_1822 %>% filter(RESPONDENT == 1) %>% select(YEAR, TOTL_SAL),
  dist_1923 %>% filter(RESPONDENT == 1) %>% select(YEAR, TOTL_SAL),
  dist_2125 %>% filter(RESPONDENT == 1) %>% select(YEAR, TOTL_SAL)
)
sermed <- ser %>%
  filter(!is.na(TOTL_SAL)) %>%
  group_by(YEAR) %>%
  summarise(median = median(TOTL_SAL), n = n(), .groups = "drop")
kable(sermed, caption = "F. Median annual salary by year — BGS", digits = 0)
```

```{r bgs-g}
# SECTION G - Occupation: map 5-digit NOC codes to NOC-2021 broad categories
# via the lookup table (nocmap). NOC-2021 codes are comparable ONLY for
# 2023-2025 (DIST_18_22 carries NOC-2016; see map ticket 05), so the
# benchmark window is just 2023 while the new-pool window is 2024-25.
g23 <- dist_1923 %>% filter(RESPONDENT == 1, YEAR == 2023)
gnew <- dist_2125 %>% filter(RESPONDENT == 1, YEAR %in% 2024:2025)
nocmap <- noc %>% select(UNIT_GROUP_CODE, BROAD_CATEGORY_ENGLISH_NAME)
# gcat(): left-join a data frame of NOC codes onto the category lookup
gcat <- function(d) {
  m <- merge(
    data.frame(NOC = d$NOC_2021),
    nocmap,
    by.x = "NOC",
    by.y = "UNIT_GROUP_CODE",
    all.x = TRUE
  )
  m$BROAD_CATEGORY_ENGLISH_NAME
}
# category counts, dropped-empty names, then shares of the category total
c23 <- table(gcat(g23))
cn <- table(gcat(gnew))
c23 <- c23[names(c23) != ""]
cn <- cn[names(cn) != ""]
# share table keyed on benchmark categories; new shares mapped by name so
# categories absent from one pool still line up
gtab <- data.frame(
  `Occupation group` = names(c23),
  Benchmark_2023 = as.numeric(pct(c23, sum(c23))),
  New_2024_25 = as.numeric(pct(cn[names(c23)], sum(cn))),
  check.names = FALSE
)
gtab <- gtab[order(-gtab$New_2024_25), ]
kable(
  gtab,
  caption = "G. NOC broad categories — BGS (NOC-2021; benchmark window 2023 only, see notes)",
  digits = 1
)
```

```{r bgs-h}
# SECTION H - Further education among respondents. TOOK_FURTH_ED: any
# positive value = continuing (1), 0 = not; NA/sentinels = missing (via
# valid()). Denominator for Continuing/Not = valid responses; Missing =
# share of all respondents.
hb <- function(d) {
  v <- d$TOOK_FURTH_ED[d$resp]
  k <- valid(v)
  c(
    Continuing = pct(sum(v[k] >= 1, na.rm = TRUE), sum(k)),
    Not = pct(sum(v[k] == 0, na.rm = TRUE), sum(k)),
    Missing = pct(sum(!k), sum(d$resp))
  )
}
htab <- data.frame(
  Metric = c("Continuing education", "Not continuing", "Missing/sentinel"),
  Benchmark = unname(hb(bgs_b)),
  New = unname(hb(bgs_n))
)
kable(htab, caption = "H. Further education — BGS", digits = 1)
```

```{r bgs-i}
# SECTION I - Satisfaction, BGS only (DIST tables). Three 4-point items:
# B03_R1 education satisfaction, E28 program relevance, C01 instruction
# quality. top-2-box = score >= 3 of 4; valid() excludes sentinels.
# eb/en are the DIST respondents defined in section D.
sat <- function(d) {
  v <- d$B03_R1
  k <- valid(v)
  c(Satisfied_top2box = pct(sum(v[k] >= 3, na.rm = TRUE), sum(k)), N = sum(k))
}
rel <- function(d) {
  v <- d$E28
  k <- valid(v)
  c(Useful_top2box = pct(sum(v[k] >= 3, na.rm = TRUE), sum(k)), N = sum(k))
}
instr <- function(d) {
  v <- d$C01
  k <- valid(v)
  c(Good_top2box = pct(sum(v[k] >= 3, na.rm = TRUE), sum(k)), N = sum(k))
}
kable(
  data.frame(
    Outcome = c(
      "Education satisfaction (B03_R1, top-2-box)",
      "Program relevance (E28, top-2-box)",
      "Instruction quality (C01, top-2-box)"
    ),
    Benchmark = c(
      paste0(sat(eb)["Satisfied_top2box"], "% (n=", sat(eb)["N"], ")"),
      paste0(rel(eb)["Useful_top2box"], "% (n=", rel(eb)["N"], ")"),
      paste0(instr(eb)["Good_top2box"], "% (n=", instr(eb)["N"], ")")
    ),
    New = c(
      paste0(sat(en)["Satisfied_top2box"], "% (n=", sat(en)["N"], ")"),
      paste0(rel(en)["Useful_top2box"], "% (n=", rel(en)["N"], ")"),
      paste0(instr(en)["Good_top2box"], "% (n=", instr(en)["N"], ")")
    )
  ),
  caption = "I. Satisfaction — BGS (4-pt scale, -1/-2 excluded)",
  digits = 1
)
```

```{r bgs-j}
# SECTION J - Program coverage: how much of the PSSM program universe is
# represented by respondents? Universe = current programs (IN_CURRENT_DATA_
# EXTRACT == "Y") with an awarded degree (ADGR) or university-transfer (UT)
# credential. BGS response tables only expose 4-digit CIP, so the universe's
# 6-digit CIPs are truncated to 4 for the join (BGS-specific; TRD/APPSO/DACSO
# use CIP6 directly).
pcur <- progs %>% filter(IN_CURRENT_DATA_EXTRACT == "Y")
univ_bgs <- pcur %>%
  filter(PRGM_CREDENTIAL %in% c("ADGR", "UT")) %>%
  distinct(PRGM_INST_CD, LCIP_CD_CIP2021)

# BGS: INST x CIP4 vs universe (CIP6 first 4)
univ_bgs4 <- univ_bgs %>%
  mutate(CIP4 = substr(LCIP_CD_CIP2021, 1, 4)) %>%
  distinct(PRGM_INST_CD, CIP4)
k4b <- bgs_b %>%
  filter(resp) %>%
  distinct(INST, CIP_CODE_4) %>%
  rename(PRGM_INST_CD = INST, CIP4 = CIP_CODE_4)
k4n <- bgs_n %>%
  filter(resp) %>%
  distinct(INST, CIP_CODE_4) %>%
  rename(PRGM_INST_CD = INST, CIP4 = CIP_CODE_4)
cover4 <- function(k) {
  nrow(inner_join(univ_bgs4, k, by = c("PRGM_INST_CD", "CIP4"))) /
    nrow(univ_bgs4) *
    100
}
cov_b <- cover4(k4b)
cov_n <- cover4(k4n)
kable(
  data.frame(
    Metric = c(
      "Universe programs (current, ADGR+UT)",
      "Programs with BGS respondents",
      "Coverage rate (%)"
    ),
    Benchmark = c(
      nrow(univ_bgs4),
      nrow(inner_join(univ_bgs4, k4b, by = c("PRGM_INST_CD", "CIP4"))),
      round(cov_b, 1)
    ),
    New = c(
      nrow(univ_bgs4),
      nrow(inner_join(univ_bgs4, k4n, by = c("PRGM_INST_CD", "CIP4"))),
      round(cov_n, 1)
    )
  ),
  caption = "J. Program coverage — BGS (INST x CIP4 vs INFOWARE_PROGRAMS current degree/UT programs)"
)
cov_delta <- cov_n - cov_b
kable(
  data.frame(
    Coverage_delta_pp = round(cov_delta, 1),
    Flag = flag(cov_delta, 5)
  ),
  caption = "J. Flag check"
)
```

```{r bgs-k}
# SECTION K - Institution-level outcomes. For every institution: respondent
# count, employment rate, and median salary (DIST only; the response tables
# carry no earnings). Median requires >= 10 non-missing salaries, else NA.
# Suppression: cells with <10 NEW-pool respondents are SUPPRESSED; 10-29 =
# "caution" (small-cell policy). Flags: employment delta > 10pp, salary
# median change > 15%, both only when the cell is not suppressed.
kinst <- function(d, distd) {
  inst <- unique(d$INST)
  # iterates institutions: loop body builds one summary row per institution
  do.call(
    rbind,
    lapply(inst, function(i) {
      s <- d %>% filter(resp, INST == i)
      es <- emp_status(s, "EMPLOYED", "UNEMPLOYED", "IN_LBR_FRC", s$resp)
      e <- pct(sum(es == "Employed", na.rm = TRUE), sum(!is.na(es)))
      sal <- distd$TOTL_SAL[distd$RESPONDENT == 1 & distd$INSTITUTION_CODE == i]
      med <- if (length(sal[!is.na(sal)]) >= 10) {
        round(median(sal, na.rm = TRUE))
      } else {
        NA
      }
      data.frame(
        Institution = i,
        Respondents = sum(s$resp),
        Employment = e,
        Median_salary = med
      )
    })
  )
}
k_b <- kinst(bgs_b, eb)
k_n <- kinst(bgs_n, en)
k_join <- merge(k_b, k_n, by = "Institution", suffixes = c("_bench", "_new"))
k_join$Suppress <- ifelse(
  k_join$Respondents_new < 10,
  "SUPPRESSED",
  ifelse(k_join$Respondents_new < 30, "caution", "")
)
k_join$Emp_flag <- ifelse(
  k_join$Respondents_new < 10,
  "",
  flag(k_join$Employment_new - k_join$Employment_bench, 10)
)
k_join$Sal_flag <- ifelse(
  k_join$Respondents_new < 10,
  "",
  flag(
    (k_join$Median_salary_new - k_join$Median_salary_bench) /
      k_join$Median_salary_bench *
      100,
    15,
    "%"
  )
)
kable(
  k_join,
  caption = "K. Institution-level outcomes — BGS (cells <10 respondents suppressed)",
  digits = 1
)
```

```{r bgs-l}
# SECTION L - Program-level employment deltas (INST x CIP4). Employment rate
# per program per pool, then flag programs whose new-pool employment differs
# from benchmark by > 10pp. Floor: >= 30 new-pool respondents (programs with
# fewer respondents are too noisy - see previous session). cur_data() = the current
# group inside summarise(). LCP4_DIGITS_NAME = CIP2021 4-digit English name
# (BGS response data has no CIP6; carried via first(), see TRD section L).
lkey <- c("INST", "CIP_CODE_4")
lb <- bgs_b %>%
  filter(resp) %>%
  group_by(INST, CIP_CODE_4) %>%
  summarise(
    nb = n(),
    LCP4_DIGITS_NAME = first(LCP4_DIGITS_NAME),
    eb = pct(
      sum(
        emp_status(cur_data(), "EMPLOYED", "UNEMPLOYED", "IN_LBR_FRC", TRUE) ==
          "Employed",
        na.rm = TRUE
      ),
      sum(
        !is.na(emp_status(
          cur_data(),
          "EMPLOYED",
          "UNEMPLOYED",
          "IN_LBR_FRC",
          TRUE
        ))
      )
    ),
    .groups = "drop"
  )
ln <- bgs_n %>%
  filter(resp) %>%
  group_by(INST, CIP_CODE_4) %>%
  summarise(
    nn = n(),
    en = pct(
      sum(
        emp_status(cur_data(), "EMPLOYED", "UNEMPLOYED", "IN_LBR_FRC", TRUE) ==
          "Employed",
        na.rm = TRUE
      ),
      sum(
        !is.na(emp_status(
          cur_data(),
          "EMPLOYED",
          "UNEMPLOYED",
          "IN_LBR_FRC",
          TRUE
        ))
      )
    ),
    .groups = "drop"
  )
lj <- merge(lb, ln, by = lkey, all = TRUE)
# name carried only by lb (bn/trd_n sources have no name col) -> single col
lj <- lj %>%
  filter(nn >= 30) %>%
  mutate(delta = en - eb) %>%
  filter(!is.na(delta) & abs(delta) > 10) %>%
  arrange(-abs(delta))
kable(
  lj[c("INST", "CIP_CODE_4", "LCP4_DIGITS_NAME", "nb", "eb", "nn", "en", "delta")],
  col.names = c("INST", "CIP4", "CIP name (EN)", "n bench", "emp % bench", "n new", "emp % new", "delta pp"),
  caption = paste0(
    "L. Program-level employment flags — BGS (INST x CIP4, floor >=30 new respondents, |delta| > 10pp). Flagged: ",
    nrow(lj)
  ),
  digits = 1
)
# low-response programs
low <- ln %>% filter(nn < 30) %>% summarise(n = n())
kable(
  low,
  caption = "L. BGS programs with <30 new-pool respondents (insufficient for program-level outcomes)"
)
```

### Verdict — BGS (draft)

```{r bgs-v}
# ----------------------------------------------------------
# Verdict (draft) - BGS: interpretation finalized per map
# previous session (resolved). Hard flag fires when the pooled
# response-rate delta exceeds 5pp; any hard flag -> REVIEW.
# Here the RR decline (-13.8pp) and the +16.1% earnings shift
# both fire. previous session confirmed: the RR drop is a real
# participation decline (frames growing,
# telephone share rebounded while RR still fell, no field-timing
# data) - NOT an artifact; the earnings shift is systemic real
# wage growth across 15/16 institutions. Both kept as hard
# flags -> REVIEW, acceptable with caveats.
# ----------------------------------------------------------
hard <- any(abs(delta_rr) > 5)
cat(
  "Hard flags: response-rate delta =",
  round(delta_rr, 1),
  "pp (|delta|>5pp:",
  hard,
  ")\n"
)
cat(
  "Verdict: REVIEW — acceptable with caveats.\n"
)
cat(
  "Hard flags: (1) pooled response rate 39.7% -> 25.9% (-13.8pp); 
(2) BGS median earnings +16.1% ($58,000 -> $67,329).\n"
)
cat(
  "Interpretation (previous session): the RR decline is a real participation 
trend, not a collection artifact — frames stable/growing (22,900 -> 
24,306 rows) and telephone share rebounded in 2024-25 (32.4%/31.3% vs 
19.5% in 2023) while RR kept falling; SURVMON all NA, so no field-timing 
evidence. The earnings shift is real systemic wage growth across 15 of 16 
institutions (+15.5% to +32.2%), not a cell-level artifact. Program-level 
employment deltas (|delta|>10pp, floor >=30 new respondents): flagged 
programs are listed in section L and treated as small-n noise, not 
blocking.",
  sep = ""
)
cat(
  "\nFollow-ups: analyst sign-off of this report; survey-team notification 
of the secular participation decline; outcomes module (02b) must apply 
non-response weights to the 2024-25 pool given the -13.8pp RR drop."
)
```

---

# TRD — Trades Graduate Survey

```{r trd-a}
# SECTION A (TRD) - same overview layout as BGS section A. TRD cycles run
# C_Outc19..23 (bench) / C_Outc24..25 (new); program key is INST x 6-digit
# LCP6_CD (not CIP4 - TRD tables carry full 6-digit CIP). Frame = all rows.
tb <- function(cond) sum(trd$resp & cond)
kable(
  data.frame(
    Metric = c(
      "Records (frame)",
      "Respondents",
      "Institutions",
      "Programs (INST x CIP6)",
      "Survey years",
      "Response rate"
    ),
    Benchmark = c(
      sum(trd$pool == pool_b),
      tb(trd$pool == pool_b),
      length(unique(trd$INST[trd$pool == pool_b])),
      nrow(distinct(trd[trd$pool == pool_b & trd$resp, c("INST", "LCP6_CD")])),
      "C_Outc19..23",
      pct(sum(trd$resp & trd$pool == pool_b), sum(trd$pool == pool_b))
    ),
    New = c(
      sum(trd$pool == pool_n),
      tb(trd$pool == pool_n),
      length(unique(trd$INST[trd$pool == pool_n])),
      nrow(distinct(trd[trd$pool == pool_n & trd$resp, c("INST", "LCP6_CD")])),
      "C_Outc24..25",
      pct(sum(trd$resp & trd$pool == pool_n), sum(trd$pool == pool_n))
    )
  ),
  caption = "A. Dataset overview — TRD"
)
```

```{r trd-b}
# SECTION B (TRD) - response rate identical logic to BGS section B; TRD
# `cycle` = cycle code parsed from SUBM_CD (C_Outc19 -> 19). d_rr = delta.
rr <- trd %>%
  group_by(cycle, pool) %>%
  summarise(
    frame = n(),
    resp = sum(resp),
    rate = pct(resp, frame),
    .groups = "drop"
  )
kable(rr, caption = "B. Response rate by cycle — TRD", digits = 1)
rr_agg <- rr %>% group_by(pool) %>% summarise(rate = pct(sum(resp), sum(frame)))
d_rr <- rr_agg$rate[rr_agg$pool == pool_n] - rr_agg$rate[rr_agg$pool == pool_b]
kable(
  data.frame(Pool = rr_agg$pool, Rate = round(rr_agg$rate, 1)),
  caption = "B. Pooled response rate — TRD"
)
kable(
  data.frame(Delta_pp = round(d_rr, 1), Flag = flag(d_rr, 5)),
  caption = "B. Flag check"
)
ggplot(rr, aes(cycle, rate, color = pool)) +
  geom_line() +
  geom_point() +
  labs(title = "TRD response rate by cycle", y = "%", x = "Cycle") +
  theme_minimal()
```

```{r trd-c}
# SECTION C (TRD) - representativeness: age distribution of the full frame
# (population proxy: TRD has NO separate eligible/cohort table - the response
# table IS the frame, unlike BGS which uses COHORT_INFO) vs respondents.
# rep_c() is reused by APPSO and DACSO with their age/institution column
# names (agev/instv). Age bands from agbins() on both sides.
rep_c <- function(d, agev, instv) {
  data.frame(
    `Age group` = c("<25", "25-34", "35+"),
    Population = c(
      # Population column: shares over rows with non-missing age (full frame)
      pct(
        sum(agbins(d[[agev]]) == "<25", na.rm = TRUE),
        sum(!is.na(d[[agev]]))
      ),
      pct(
        sum(agbins(d[[agev]]) == "25-34", na.rm = TRUE),
        sum(!is.na(d[[agev]]))
      ),
      pct(sum(agbins(d[[agev]]) == "35+", na.rm = TRUE), sum(!is.na(d[[agev]])))
    ),
    Respondents = c(
      # Respondents column: same bands but subset to resp == TRUE; d$resp is
      # the respondent flag carried by the combined datasets
      pct(
        sum(agbins(d[[agev]][d$resp]) == "<25", na.rm = TRUE),
        sum(!is.na(d[[agev]][d$resp]))
      ),
      pct(
        sum(agbins(d[[agev]][d$resp]) == "25-34", na.rm = TRUE),
        sum(!is.na(d[[agev]][d$resp]))
      ),
      pct(
        sum(agbins(d[[agev]][d$resp]) == "35+", na.rm = TRUE),
        sum(!is.na(d[[agev]][d$resp]))
      )
    )
  )
}
kable(
  rep_c(trd, "TRD_AGE_AT_SURVEY", "INST"),
  caption = "C. Age — TRD frame vs respondents (new pool)",
  digits = 1
)
```

```{r trd-d}
# SECTION D (TRD) - missingness among respondents, same logic as BGS; note
# TRD column naming (TRD_LABR_EMPLOYED, NOC_CD).
miss_trd <- data.frame(
  Variable = c("Employment status", "Occupation (NOC_CD)"),
  Benchmark = c(
    pct(sum(is.na(trd_b$TRD_LABR_EMPLOYED) & trd_b$resp), sum(trd_b$resp)),
    pct(
      sum((is.na(trd_b$NOC_CD) | trd_b$NOC_CD == "XXXXX") & trd_b$resp),
      sum(trd_b$resp)
    )
  ),
  New = c(
    pct(sum(is.na(trd_n$TRD_LABR_EMPLOYED) & trd_n$resp), sum(trd_n$resp)),
    pct(
      sum((is.na(trd_n$NOC_CD) | trd_n$NOC_CD == "XXXXX") & trd_n$resp),
      sum(trd_n$resp)
    )
  )
)
kable(
  miss_trd,
  caption = "D. Missingness among TRD respondents (%)",
  digits = 1
)
```

```{r trd-e}
# SECTION E (TRD) - employment status via emp_status() with TRD column names.
# No earnings/satisfaction for TRD (BGS-only via DIST), so no F/I sections.
e1 <- emp_status(
  trd_b,
  "TRD_LABR_EMPLOYED",
  "TRD_LABR_UNEMPLOYED",
  "TRD_LABR_IN_LABOUR_MARKET",
  trd_b$resp
)
e2 <- emp_status(
  trd_n,
  "TRD_LABR_EMPLOYED",
  "TRD_LABR_UNEMPLOYED",
  "TRD_LABR_IN_LABOUR_MARKET",
  trd_n$resp
)
etab <- data.frame(
  Status = c("Employed", "Unemployed", "Not in labour force"),
  Benchmark = c(
    pct(sum(e1 == "Employed", na.rm = TRUE), sum(!is.na(e1))),
    pct(sum(e1 == "Unemployed", na.rm = TRUE), sum(!is.na(e1))),
    pct(sum(e1 == "Not in labour force", na.rm = TRUE), sum(!is.na(e1)))
  ),
  New = c(
    pct(sum(e2 == "Employed", na.rm = TRUE), sum(!is.na(e2))),
    pct(sum(e2 == "Unemployed", na.rm = TRUE), sum(!is.na(e2))),
    pct(sum(e2 == "Not in labour force", na.rm = TRUE), sum(!is.na(e2)))
  )
)
etab$Status <- factor(
  etab$Status,
  levels = c("Employed", "Unemployed", "Not in labour force")
)
kable(
  etab,
  caption = "E. Employment status — TRD (valid-status base)",
  digits = 1
)
epool <- etab %>%
  pivot_longer(-Status, names_to = "Pool", values_to = "Share") %>%
  mutate(Pool = factor(Pool, levels = c("Benchmark", "New")))
ggplot(epool, aes(Pool, Share, fill = Status)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title = "E. Employment status — TRD benchmark vs new",
    y = "Share",
    x = NULL
  ) +
  theme_minimal()
e_d <- etab$New[1] - etab$Benchmark[1]
kable(
  data.frame(Employed_delta_pp = round(e_d, 1), Flag = flag(e_d, 10)),
  caption = "E. Flag check"
)
```

```{r trd-g}
# SECTION G (TRD) - occupation: NOC_CD (NOC-2021) from the response table
# (TRD has no DIST equivalent), mapped via nocmap. Both pools are NOC-2021,
# so the full benchmark window is comparable (unlike BGS/DACSO - see section
# G notes). Merges parallel the BGS gcat() approach.
g1 <- merge(
  data.frame(NOC = trd_b$NOC_CD[trd_b$resp]),
  nocmap,
  by.x = "NOC",
  by.y = "UNIT_GROUP_CODE",
  all.x = TRUE
)
g2 <- merge(
  data.frame(NOC = trd_n$NOC_CD[trd_n$resp]),
  nocmap,
  by.x = "NOC",
  by.y = "UNIT_GROUP_CODE",
  all.x = TRUE
)
c1 <- table(g1$BROAD_CATEGORY_ENGLISH_NAME)
c2 <- table(g2$BROAD_CATEGORY_ENGLISH_NAME)
c1 <- c1[names(c1) != ""]
c2 <- c2[names(c2) != ""]
gt <- data.frame(
  `Occupation group` = names(c1),
  Benchmark = as.numeric(pct(c1, sum(c1))),
  New = as.numeric(pct(c2[names(c1)], sum(c2))),
  check.names = FALSE
)
gt <- gt[order(-gt$New), ]
kable(
  gt,
  caption = "G. NOC broad categories — TRD (NOC-2021 both pools)",
  digits = 1
)
```

```{r trd-k}
# SECTION K (TRD) - institution-level employment, same pattern as BGS section
# K but without salary (no DIST for TRD): respondents + employment per
# institution; <10 new-pool respondents suppressed, 10-29 caution; delta
# flag > 10pp.
kinst_t <- function(d) {
  inst <- unique(d$INST)
  do.call(
    rbind,
    lapply(inst, function(i) {
      s <- d %>% filter(resp, INST == i)
      es <- emp_status(
        s,
        "TRD_LABR_EMPLOYED",
        "TRD_LABR_UNEMPLOYED",
        "TRD_LABR_IN_LABOUR_MARKET",
        s$resp
      )
      data.frame(
        Institution = i,
        Respondents = sum(s$resp),
        Employment = pct(sum(es == "Employed", na.rm = TRUE), sum(!is.na(es)))
      )
    })
  )
}
k_t <- merge(
  kinst_t(trd_b),
  kinst_t(trd_n),
  by = "Institution",
  suffixes = c("_bench", "_new")
)
k_t$Suppress <- ifelse(
  k_t$Respondents_new < 10,
  "SUPPRESSED",
  ifelse(k_t$Respondents_new < 30, "caution", "")
)
k_t$Emp_flag <- ifelse(
  k_t$Respondents_new < 10,
  "",
  flag(k_t$Employment_new - k_t$Employment_bench, 10)
)
kable(
  k_t,
  caption = "K. Institution-level employment — TRD (cells <10 suppressed)",
  digits = 1
)
```

```{r trd-l}
# SECTION L (TRD) - program-level employment deltas at INST x 6-digit CIP
# (LCP6_CD), floor >= 30 new respondents, |delta| > 10pp flagged. Logic as
# BGS section L. LCP6_DIGITS_NAME = CIP2021 English name joined in at load
# time (carried through with first() since summarise() drops non-group cols;
# a few survey codes sit outside the CIP2021 classification -> blank).
lb <- trd_b %>%
  filter(resp) %>%
  group_by(INST, LCP6_CD) %>%
  summarise(
    nb = n(),
    LCP6_DIGITS_NAME = first(LCP6_DIGITS_NAME),
    eb = pct(
      sum(
        emp_status(
          cur_data(),
          "TRD_LABR_EMPLOYED",
          "TRD_LABR_UNEMPLOYED",
          "TRD_LABR_IN_LABOUR_MARKET",
          TRUE
        ) ==
          "Employed",
        na.rm = TRUE
      ),
      sum(
        !is.na(emp_status(
          cur_data(),
          "TRD_LABR_EMPLOYED",
          "TRD_LABR_UNEMPLOYED",
          "TRD_LABR_IN_LABOUR_MARKET",
          TRUE
        ))
      )
    ),
    .groups = "drop"
  )
ln <- trd_n %>%
  filter(resp) %>%
  group_by(INST, LCP6_CD) %>%
  summarise(
    nn = n(),
    en = pct(
      sum(
        emp_status(
          cur_data(),
          "TRD_LABR_EMPLOYED",
          "TRD_LABR_UNEMPLOYED",
          "TRD_LABR_IN_LABOUR_MARKET",
          TRUE
        ) ==
          "Employed",
        na.rm = TRUE
      ),
      sum(
        !is.na(emp_status(
          cur_data(),
          "TRD_LABR_EMPLOYED",
          "TRD_LABR_UNEMPLOYED",
          "TRD_LABR_IN_LABOUR_MARKET",
          TRUE
        ))
      )
    ),
    .groups = "drop"
  )
lj <- merge(lb, ln, by = c("INST", "LCP6_CD"), all = TRUE)
# name carried only by lb (ln sources have no name col) -> single col
lj <- lj %>%
  filter(nn >= 30) %>%
  mutate(delta = en - eb) %>%
  filter(!is.na(delta) & abs(delta) > 10) %>%
  arrange(-abs(delta))
kable(
  lj[c("INST", "LCP6_CD", "LCP6_DIGITS_NAME", "nb", "eb", "nn", "en", "delta")],
  col.names = c("INST", "CIP6", "CIP name (EN)", "n bench", "emp % bench", "n new", "emp % new", "delta pp"),
  caption = paste0(
    "L. Program-level employment flags — TRD (INST x CIP6, floor >=30, |delta| > 10pp). Flagged: ",
    nrow(lj)
  ),
  digits = 1
)
```

### Verdict — TRD (draft)

```{r trd-v}
# ----------------------------------------------------------
# Verdict (draft) - TRD: interpretation finalized per map
# previous session (resolved). Hard
# flag from d_rr (response-rate delta); -10.6pp fires ->
# REVIEW, acceptable with caveats. Employment stable; only 3
# program-level flags, none blocking.
# ----------------------------------------------------------
hard <- any(abs(d_rr) > 5)
cat(
  "Hard flags: response-rate delta =",
  round(d_rr, 1),
  "pp (|delta|>5pp:",
  hard,
  ")\n"
)
cat(
  "Verdict: REVIEW — acceptable with caveats.\n"
)
cat(
  "Hard flag: pooled response rate 48.5% -> 37.9% (-10.6pp); employment 
rate stable (83.2% vs 82.8%). Three program-level employment deltas 
(|delta|>10pp, floor >=30) flagged — listed in section L as small-n 
noise, not blocking.",
  sep = ""
)
cat(
  "\nFollow-ups: analyst sign-off; survey-team notification of participation 
decline; outcomes module (02b) non-response weighting for the 2024-25 pool."
)
```

---

# APPSO — Apprenticeship Survey

```{r appso-a}
# SECTION A (APPSO) - overview, same layout as BGS/TRD section A; cycles and
# frame semantics identical (frame = all rows), programs at INST x CIP6.
ab <- function(cond) sum(app$resp & cond)
kable(
  data.frame(
    Metric = c(
      "Records (frame)",
      "Respondents",
      "Institutions",
      "Programs (INST x CIP6)",
      "Survey years",
      "Response rate"
    ),
    Benchmark = c(
      sum(app$pool == pool_b),
      ab(app$pool == pool_b),
      length(unique(app$INST[app$pool == pool_b])),
      nrow(distinct(app[app$pool == pool_b & app$resp, c("INST", "LCP6_CD")])),
      "C_Outc19..23",
      pct(sum(app$resp & app$pool == pool_b), sum(app$pool == pool_b))
    ),
    New = c(
      sum(app$pool == pool_n),
      ab(app$pool == pool_n),
      length(unique(app$INST[app$pool == pool_n])),
      nrow(distinct(app[app$pool == pool_n & app$resp, c("INST", "LCP6_CD")])),
      "C_Outc24..25",
      pct(sum(app$resp & app$pool == pool_n), sum(app$pool == pool_n))
    )
  ),
  caption = "A. Dataset overview — APPSO"
)
```

```{r appso-b}
# SECTION B (APPSO) - response rate, identical logic to TRD section B.
rr <- app %>%
  group_by(cycle, pool) %>%
  summarise(
    frame = n(),
    resp = sum(resp),
    rate = pct(resp, frame),
    .groups = "drop"
  )
kable(rr, caption = "B. Response rate by cycle — APPSO", digits = 1)
rr_agg <- rr %>% group_by(pool) %>% summarise(rate = pct(sum(resp), sum(frame)))
d_rr <- rr_agg$rate[rr_agg$pool == pool_n] - rr_agg$rate[rr_agg$pool == pool_b]
kable(
  data.frame(Pool = rr_agg$pool, Rate = round(rr_agg$rate, 1)),
  caption = "B. Pooled response rate — APPSO"
)
kable(
  data.frame(Delta_pp = round(d_rr, 1), Flag = flag(d_rr, 5)),
  caption = "B. Flag check"
)
ggplot(rr, aes(cycle, rate, color = pool)) +
  geom_line() +
  geom_point() +
  labs(title = "APPSO response rate by cycle", y = "%", x = "Cycle") +
  theme_minimal()
```

```{r appso-c}
# SECTION C (APPSO) - age representativeness, reuses rep_c() from TRD section
# C with APPSO's age column (APP_AGE_AT_SURVEY).
kable(
  rep_c(app, "APP_AGE_AT_SURVEY", "INST"),
  caption = "C. Age — APPSO frame vs respondents (new pool)",
  digits = 1
)
```

```{r appso-d}
# SECTION D (APPSO) - missingness among respondents; same logic as TRD.
# Note the occupation row: NOC_CD is largely UNCOLLECTED in APPSO (~8% of
# respondents carry a code - collection gap, see map tickets 02/07).
miss_app <- data.frame(
  Variable = c("Employment status", "Occupation (NOC_CD)"),
  Benchmark = c(
    pct(sum(is.na(app_b$APP_LABR_EMPLOYED) & app_b$resp), sum(app_b$resp)),
    pct(
      sum((is.na(app_b$NOC_CD) | app_b$NOC_CD == "XXXXX") & app_b$resp),
      sum(app_b$resp)
    )
  ),
  New = c(
    pct(sum(is.na(app_n$APP_LABR_EMPLOYED) & app_n$resp), sum(app_n$resp)),
    pct(
      sum((is.na(app_n$NOC_CD) | app_n$NOC_CD == "XXXXX") & app_n$resp),
      sum(app_n$resp)
    )
  )
)
kable(
  miss_app,
  caption = "D. Missingness among APPSO respondents (%) — occupation largely uncollected",
  digits = 1
)
```

```{r appso-e}
# SECTION E (APPSO) - employment status via emp_status() with APPSO column
# names (APP_LABR_*). No earnings/satisfaction for APPSO.
e1 <- emp_status(
  app_b,
  "APP_LABR_EMPLOYED",
  "APP_LABR_UNEMPLOYED",
  "APP_LABR_IN_LABOUR_MARKET",
  app_b$resp
)
e2 <- emp_status(
  app_n,
  "APP_LABR_EMPLOYED",
  "APP_LABR_UNEMPLOYED",
  "APP_LABR_IN_LABOUR_MARKET",
  app_n$resp
)
etab <- data.frame(
  Status = c("Employed", "Unemployed", "Not in labour force"),
  Benchmark = c(
    pct(sum(e1 == "Employed", na.rm = TRUE), sum(!is.na(e1))),
    pct(sum(e1 == "Unemployed", na.rm = TRUE), sum(!is.na(e1))),
    pct(sum(e1 == "Not in labour force", na.rm = TRUE), sum(!is.na(e1)))
  ),
  New = c(
    pct(sum(e2 == "Employed", na.rm = TRUE), sum(!is.na(e2))),
    pct(sum(e2 == "Unemployed", na.rm = TRUE), sum(!is.na(e2))),
    pct(sum(e2 == "Not in labour force", na.rm = TRUE), sum(!is.na(e2)))
  )
)
etab$Status <- factor(
  etab$Status,
  levels = c("Employed", "Unemployed", "Not in labour force")
)
kable(
  etab,
  caption = "E. Employment status — APPSO (valid-status base)",
  digits = 1
)
epool <- etab %>%
  pivot_longer(-Status, names_to = "Pool", values_to = "Share") %>%
  mutate(Pool = factor(Pool, levels = c("Benchmark", "New")))
ggplot(epool, aes(Pool, Share, fill = Status)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title = "E. Employment status — APPSO benchmark vs new",
    y = "Share",
    x = NULL
  ) +
  theme_minimal()
e_d <- etab$New[1] - etab$Benchmark[1]
kable(
  data.frame(Employed_delta_pp = round(e_d, 1), Flag = flag(e_d, 10)),
  caption = "E. Flag check"
)
```

```{r appso-g}
# SECTION G (APPSO) - occupation via NOC_CD. CAVEAT: APPSO occupation is
# largely uncollected (~8% of respondents carry codes) - results here are
# illustrative only (see map tickets 02/07).
g1 <- merge(
  data.frame(NOC = app_b$NOC_CD[app_b$resp]),
  nocmap,
  by.x = "NOC",
  by.y = "UNIT_GROUP_CODE",
  all.x = TRUE
)
g2 <- merge(
  data.frame(NOC = app_n$NOC_CD[app_n$resp]),
  nocmap,
  by.x = "NOC",
  by.y = "UNIT_GROUP_CODE",
  all.x = TRUE
)
c1 <- table(g1$BROAD_CATEGORY_ENGLISH_NAME)
c2 <- table(g2$BROAD_CATEGORY_ENGLISH_NAME)
c1 <- c1[names(c1) != ""]
c2 <- c2[names(c2) != ""]
gt <- data.frame(
  `Occupation group` = names(c1),
  Benchmark = as.numeric(pct(c1, sum(c1))),
  New = as.numeric(pct(c2[names(c1)], sum(c2))),
  check.names = FALSE
)
gt <- gt[order(-gt$New), ]
kable(
  gt,
  caption = "G. NOC broad categories — APPSO (low response coverage: ~8% of respondents have codes; interpret with caution)",
  digits = 1
)
```

```{r appso-k}
# SECTION K (APPSO) - institution-level employment, same pattern as TRD
# section K (no salary for APPSO).
kinst_a <- function(d) {
  inst <- unique(d$INST)
  do.call(
    rbind,
    lapply(inst, function(i) {
      s <- d %>% filter(resp, INST == i)
      es <- emp_status(
        s,
        "APP_LABR_EMPLOYED",
        "APP_LABR_UNEMPLOYED",
        "APP_LABR_IN_LABOUR_MARKET",
        s$resp
      )
      data.frame(
        Institution = i,
        Respondents = sum(s$resp),
        Employment = pct(sum(es == "Employed", na.rm = TRUE), sum(!is.na(es)))
      )
    })
  )
}
k_a <- merge(
  kinst_a(app_b),
  kinst_a(app_n),
  by = "Institution",
  suffixes = c("_bench", "_new")
)
k_a$Suppress <- ifelse(
  k_a$Respondents_new < 10,
  "SUPPRESSED",
  ifelse(k_a$Respondents_new < 30, "caution", "")
)
k_a$Emp_flag <- ifelse(
  k_a$Respondents_new < 10,
  "",
  flag(k_a$Employment_new - k_a$Employment_bench, 10)
)
kable(
  k_a,
  caption = "K. Institution-level employment — APPSO (cells <10 suppressed)",
  digits = 1
)
```

```{r appso-l}
# SECTION L (APPSO) - program-level employment deltas at INST x CIP6, floor
# >= 30, |delta| > 10pp. Logic as BGS/TRD section L. LCP6_DIGITS_NAME =
# CIP2021 English name (carried via first(); see TRD section L).
lb <- app_b %>%
  filter(resp) %>%
  group_by(INST, LCP6_CD) %>%
  summarise(
    nb = n(),
    LCP6_DIGITS_NAME = first(LCP6_DIGITS_NAME),
    eb = pct(
      sum(
        emp_status(
          cur_data(),
          "APP_LABR_EMPLOYED",
          "APP_LABR_UNEMPLOYED",
          "APP_LABR_IN_LABOUR_MARKET",
          TRUE
        ) ==
          "Employed",
        na.rm = TRUE
      ),
      sum(
        !is.na(emp_status(
          cur_data(),
          "APP_LABR_EMPLOYED",
          "APP_LABR_UNEMPLOYED",
          "APP_LABR_IN_LABOUR_MARKET",
          TRUE
        ))
      )
    ),
    .groups = "drop"
  )
ln <- app_n %>%
  filter(resp) %>%
  group_by(INST, LCP6_CD) %>%
  summarise(
    nn = n(),
    en = pct(
      sum(
        emp_status(
          cur_data(),
          "APP_LABR_EMPLOYED",
          "APP_LABR_UNEMPLOYED",
          "APP_LABR_IN_LABOUR_MARKET",
          TRUE
        ) ==
          "Employed",
        na.rm = TRUE
      ),
      sum(
        !is.na(emp_status(
          cur_data(),
          "APP_LABR_EMPLOYED",
          "APP_LABR_UNEMPLOYED",
          "APP_LABR_IN_LABOUR_MARKET",
          TRUE
        ))
      )
    ),
    .groups = "drop"
  )
lj <- merge(lb, ln, by = c("INST", "LCP6_CD"), all = TRUE)
# name carried only by lb (ln sources have no name col) -> single col
lj <- lj %>%
  filter(nn >= 30) %>%
  mutate(delta = en - eb) %>%
  filter(!is.na(delta) & abs(delta) > 10) %>%
  arrange(-abs(delta))
kable(
  lj[c("INST", "LCP6_CD", "LCP6_DIGITS_NAME", "nb", "eb", "nn", "en", "delta")],
  col.names = c("INST", "CIP6", "CIP name (EN)", "n bench", "emp % bench", "n new", "emp % new", "delta pp"),
  caption = paste0(
    "L. Program-level employment flags — APPSO (INST x CIP6, floor >=30, |delta| > 10pp). Flagged: ",
    nrow(lj)
  ),
  digits = 1
)
```

### Verdict — APPSO (draft)

```{r appso-v}
# ----------------------------------------------------------
# Verdict (draft) - APPSO: interpretation finalized per map
# previous session (resolved). Hard
# flag from d_rr; -12.4pp fires. Occupation NOC is only ~8%
# coverage — previous session accepted this with caution as a
# collection gap; occupation results are NOT representative
# of the full APPSO cohort. Two program-level flags, none
# blocking. -> REVIEW, acceptable with caveats.
# ----------------------------------------------------------
hard <- any(abs(d_rr) > 5)
cat(
  "Hard flags: response-rate delta =",
  round(d_rr, 1),
  "pp (|delta|>5pp:",
  hard,
  ")\n"
)
cat(
  "Verdict: REVIEW — acceptable with caveats.\n"
)
cat(
  "Hard flag: pooled response rate 49.6% -> 37.2% (-12.4pp). Occupation 
(NOC) data accepted with caution — only ~8% coverage, a documented collection 
gap; occupation-section results are not representative of the full APPSO 
cohort. Two program-level employment deltas (|delta|>10pp, floor >=30) 
flagged — listed in section L as small-n noise, not blocking.",
  sep = ""
)
cat(
  "\nFollow-ups: analyst sign-off; survey-team notification of participation 
decline; work with the survey team on occupation-collection completeness 
for the next cycle; outcomes module (02b) non-response weighting for the 
2024-25 pool."
)
```

---

# DACSO — Diploma / Associate / Certificate Survey

```{r dacso-a}
# SECTION A (DACSO) - overview, same layout; note DACSO prefixes its columns
# with COCI_ (COCI_INST_CD, COCI_SUBM_CD). Frame = all rows.
db2 <- function(cond) sum(dac$resp & cond)
kable(
  data.frame(
    Metric = c(
      "Records (frame)",
      "Respondents",
      "Institutions",
      "Programs (INST x CIP6)",
      "Survey years",
      "Response rate"
    ),
    Benchmark = c(
      sum(dac$pool == pool_b),
      db2(dac$pool == pool_b),
      length(unique(dac$COCI_INST_CD[dac$pool == pool_b])),
      nrow(distinct(dac[
        dac$pool == pool_b & dac$resp,
        c("COCI_INST_CD", "LCP6_CD")
      ])),
      "C_Outc19..23",
      pct(sum(dac$resp & dac$pool == pool_b), sum(dac$pool == pool_b))
    ),
    New = c(
      sum(dac$pool == pool_n),
      db2(dac$pool == pool_n),
      length(unique(dac$COCI_INST_CD[dac$pool == pool_n])),
      nrow(distinct(dac[
        dac$pool == pool_n & dac$resp,
        c("COCI_INST_CD", "LCP6_CD")
      ])),
      "C_Outc24..25",
      pct(sum(dac$resp & dac$pool == pool_n), sum(dac$pool == pool_n))
    )
  ),
  caption = "A. Dataset overview — DACSO"
)
```

```{r dacso-b}
# SECTION B (DACSO) - response rate, identical logic to TRD section B.
# DACSO respondent flag is RESPONDENT == 1 with NULL for non-respondents
# (never 0 - handled in the load script's resp()).
rr <- dac %>%
  group_by(cycle, pool) %>%
  summarise(
    frame = n(),
    resp = sum(resp),
    rate = pct(resp, frame),
    .groups = "drop"
  )
kable(rr, caption = "B. Response rate by cycle — DACSO", digits = 1)
rr_agg <- rr %>% group_by(pool) %>% summarise(rate = pct(sum(resp), sum(frame)))
d_rr <- rr_agg$rate[rr_agg$pool == pool_n] - rr_agg$rate[rr_agg$pool == pool_b]
kable(
  data.frame(Pool = rr_agg$pool, Rate = round(rr_agg$rate, 1)),
  caption = "B. Pooled response rate — DACSO"
)
kable(
  data.frame(Delta_pp = round(d_rr, 1), Flag = flag(d_rr, 5)),
  caption = "B. Flag check"
)
ggplot(rr, aes(cycle, rate, color = pool)) +
  geom_line() +
  geom_point() +
  labs(title = "DACSO response rate by cycle", y = "%", x = "Cycle") +
  theme_minimal()
```

```{r dacso-c}
# SECTION C (DACSO) - age representativeness, reuses rep_c() with DACSO's
# age/institution column names (COCI_AGE_AT_SURVEY / COCI_INST_CD).
kable(
  rep_c(dac, "COCI_AGE_AT_SURVEY", "COCI_INST_CD"),
  caption = "C. Age — DACSO frame vs respondents (new pool)",
  digits = 1
)
```

```{r dacso-d}
# SECTION D (DACSO) - missingness among respondents; LABR_* columns, NOC in
# LABR_OCCUPATION_LNOC_CD ("XXXXX" = unknown).
miss_dac <- data.frame(
  Variable = c("Employment status", "Occupation (LNOC)"),
  Benchmark = c(
    pct(sum(is.na(dac_b$LABR_EMPLOYED) & dac_b$resp), sum(dac_b$resp)),
    pct(
      sum(
        (is.na(dac_b$LABR_OCCUPATION_LNOC_CD) |
          dac_b$LABR_OCCUPATION_LNOC_CD == "XXXXX") &
          dac_b$resp
      ),
      sum(dac_b$resp)
    )
  ),
  New = c(
    pct(sum(is.na(dac_n$LABR_EMPLOYED) & dac_n$resp), sum(dac_n$resp)),
    pct(
      sum(
        (is.na(dac_n$LABR_OCCUPATION_LNOC_CD) |
          dac_n$LABR_OCCUPATION_LNOC_CD == "XXXXX") &
          dac_n$resp
      ),
      sum(dac_n$resp)
    )
  )
)
kable(
  miss_dac,
  caption = "D. Missingness among DACSO respondents (%)",
  digits = 1
)
```

```{r dacso-e}
# SECTION E (DACSO) - employment status via emp_status() with DACSO column
# names (LABR_* - no prefix. NOTE: unlike TRD/APPSO, no COCI_ prefix on the
# labour-market columns).
e1 <- emp_status(
  dac_b,
  "LABR_EMPLOYED",
  "LABR_UNEMPLOYED",
  "LABR_IN_LABOUR_MARKET",
  dac_b$resp
)
e2 <- emp_status(
  dac_n,
  "LABR_EMPLOYED",
  "LABR_UNEMPLOYED",
  "LABR_IN_LABOUR_MARKET",
  dac_n$resp
)
etab <- data.frame(
  Status = c("Employed", "Unemployed", "Not in labour force"),
  Benchmark = c(
    pct(sum(e1 == "Employed", na.rm = TRUE), sum(!is.na(e1))),
    pct(sum(e1 == "Unemployed", na.rm = TRUE), sum(!is.na(e1))),
    pct(sum(e1 == "Not in labour force", na.rm = TRUE), sum(!is.na(e1)))
  ),
  New = c(
    pct(sum(e2 == "Employed", na.rm = TRUE), sum(!is.na(e2))),
    pct(sum(e2 == "Unemployed", na.rm = TRUE), sum(!is.na(e2))),
    pct(sum(e2 == "Not in labour force", na.rm = TRUE), sum(!is.na(e2)))
  )
)
etab$Status <- factor(
  etab$Status,
  levels = c("Employed", "Unemployed", "Not in labour force")
)
kable(
  etab,
  caption = "E. Employment status — DACSO (valid-status base)",
  digits = 1
)
epool <- etab %>%
  pivot_longer(-Status, names_to = "Pool", values_to = "Share") %>%
  mutate(Pool = factor(Pool, levels = c("Benchmark", "New")))
ggplot(epool, aes(Pool, Share, fill = Status)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title = "E. Employment status — DACSO benchmark vs new",
    y = "Share",
    x = NULL
  ) +
  theme_minimal()
e_d <- etab$New[1] - etab$Benchmark[1]
kable(
  data.frame(Employed_delta_pp = round(e_d, 1), Flag = flag(e_d, 10)),
  caption = "E. Flag check"
)
```

```{r dacso-g}
# SECTION G (DACSO) - occupation via LABR_OCCUPATION_LNOC_CD (NOC-2021).
# Like BGS, the benchmark occupation window is 2023 only (earlier benchmark
# cycles used NOC-2016, not comparable).
g1 <- merge(
  data.frame(NOC = dac_b$LABR_OCCUPATION_LNOC_CD[dac_b$resp]),
  nocmap,
  by.x = "NOC",
  by.y = "UNIT_GROUP_CODE",
  all.x = TRUE
)
g2 <- merge(
  data.frame(NOC = dac_n$LABR_OCCUPATION_LNOC_CD[dac_n$resp]),
  nocmap,
  by.x = "NOC",
  by.y = "UNIT_GROUP_CODE",
  all.x = TRUE
)
c1 <- table(g1$BROAD_CATEGORY_ENGLISH_NAME)
c2 <- table(g2$BROAD_CATEGORY_ENGLISH_NAME)
c1 <- c1[names(c1) != ""]
c2 <- c2[names(c2) != ""]
gt <- data.frame(
  `Occupation group` = names(c1),
  Benchmark = as.numeric(pct(c1, sum(c1))),
  New = as.numeric(pct(c2[names(c1)], sum(c2))),
  check.names = FALSE
)
gt <- gt[order(-gt$New), ]
kable(
  gt,
  caption = "G. NOC broad categories — DACSO (NOC-2021; benchmark window 2023 only)",
  digits = 1
)
```

```{r dacso-h}
# SECTION H (DACSO) - further education. The stepa response table's PFST_*
# columns are UNPOPULATED - real answers live in the separate "clean short
# response" table (outc), which has MULTIPLE rows per respondent key:
# dedupe keeping the row with a non-NA answer (arrange + distinct).
# resp = RESPONDENT == 1 (NULL = non-respondent). Continuing = 1, Not = 0,
# Missing = share of all respondents (see map tickets 05/07).
h_dac <- function(outc) {
  o <- outc %>%
    arrange(desc(!is.na(PFST_FURSTDY_INCL_STILL_ATTD))) %>%
    distinct(STQU_ID, SUBM_CD, .keep_all = TRUE)
  o <- o %>% mutate(resp = !is.na(RESPONDENT) & RESPONDENT == 1)
  v <- o$PFST_FURSTDY_INCL_STILL_ATTD[o$resp]
  k <- valid(v)
  c(
    Continuing = pct(sum(v[k] == 1, na.rm = TRUE), sum(k)),
    Not = pct(sum(v[k] == 0, na.rm = TRUE), sum(k)),
    Missing = pct(sum(!k), sum(o$resp))
  )
}
kable(
  data.frame(
    Metric = c("Continuing education", "Not continuing", "Missing/NA"),
    Benchmark = unname(h_dac(outc_b)),
    New = unname(h_dac(outc_n))
  ),
  caption = "H. Further education — DACSO (via clean-short-response table; deduped per respondent key)"
)
```

```{r dacso-k}
# SECTION K (DACSO) - institution-level employment, same pattern as TRD
# section K; DACSO institution column is COCI_INST_CD.
kinst_d <- function(d) {
  inst <- unique(d$COCI_INST_CD)
  do.call(
    rbind,
    lapply(inst, function(i) {
      s <- d %>% filter(resp, COCI_INST_CD == i)
      es <- emp_status(
        s,
        "LABR_EMPLOYED",
        "LABR_UNEMPLOYED",
        "LABR_IN_LABOUR_MARKET",
        s$resp
      )
      data.frame(
        Institution = i,
        Respondents = sum(s$resp),
        Employment = pct(sum(es == "Employed", na.rm = TRUE), sum(!is.na(es)))
      )
    })
  )
}
k_d <- merge(
  kinst_d(dac_b),
  kinst_d(dac_n),
  by = "Institution",
  suffixes = c("_bench", "_new")
)
k_d$Suppress <- ifelse(
  k_d$Respondents_new < 10,
  "SUPPRESSED",
  ifelse(k_d$Respondents_new < 30, "caution", "")
)
k_d$Emp_flag <- ifelse(
  k_d$Respondents_new < 10,
  "",
  flag(k_d$Employment_new - k_d$Employment_bench, 10)
)
kable(
  k_d,
  caption = "K. Institution-level employment — DACSO (cells <10 suppressed)",
  digits = 1
)
```

```{r dacso-l}
# SECTION L (DACSO) - program-level employment deltas at COCI_INST_CD x CIP6,
# floor >= 30, |delta| > 10pp. Logic as BGS/TRD section L. LCP6_DIGITS_NAME =
# CIP2021 English name (carried via first(); see TRD section L).
lb <- dac_b %>%
  filter(resp) %>%
  group_by(COCI_INST_CD, LCP6_CD) %>%
  summarise(
    nb = n(),
    LCP6_DIGITS_NAME = first(LCP6_DIGITS_NAME),
    eb = pct(
      sum(
        emp_status(
          cur_data(),
          "LABR_EMPLOYED",
          "LABR_UNEMPLOYED",
          "LABR_IN_LABOUR_MARKET",
          TRUE
        ) ==
          "Employed",
        na.rm = TRUE
      ),
      sum(
        !is.na(emp_status(
          cur_data(),
          "LABR_EMPLOYED",
          "LABR_UNEMPLOYED",
          "LABR_IN_LABOUR_MARKET",
          TRUE
        ))
      )
    ),
    .groups = "drop"
  )
ln <- dac_n %>%
  filter(resp) %>%
  group_by(COCI_INST_CD, LCP6_CD) %>%
  summarise(
    nn = n(),
    en = pct(
      sum(
        emp_status(
          cur_data(),
          "LABR_EMPLOYED",
          "LABR_UNEMPLOYED",
          "LABR_IN_LABOUR_MARKET",
          TRUE
        ) ==
          "Employed",
        na.rm = TRUE
      ),
      sum(
        !is.na(emp_status(
          cur_data(),
          "LABR_EMPLOYED",
          "LABR_UNEMPLOYED",
          "LABR_IN_LABOUR_MARKET",
          TRUE
        ))
      )
    ),
    .groups = "drop"
  )
lj <- merge(lb, ln, by = c("COCI_INST_CD", "LCP6_CD"), all = TRUE)
# name carried only by lb (ln sources have no name col) -> single col
lj <- lj %>%
  filter(nn >= 30) %>%
  mutate(delta = en - eb) %>%
  filter(!is.na(delta) & abs(delta) > 10) %>%
  arrange(-abs(delta))
kable(
  lj[c("COCI_INST_CD", "LCP6_CD", "LCP6_DIGITS_NAME", "nb", "eb", "nn", "en", "delta")],
  col.names = c("INST", "CIP6", "CIP name (EN)", "n bench", "emp % bench", "n new", "emp % new", "delta pp"),
  caption = paste0(
    "L. Program-level employment flags — DACSO (INST x CIP6, floor >=30, |delta| > 10pp). Flagged: ",
    nrow(lj)
  ),
  digits = 1
)
```

### Verdict — DACSO (draft)

```{r dacso-v}
# ----------------------------------------------------------
# Verdict (draft) - DACSO: interpretation finalized per map
# previous session (resolved). Hard
# flag from d_rr; -12.8pp fires. 26 program-level flags — all
# small-n noise, non-blocking per previous session. -> REVIEW,
# acceptable with caveats.
# ----------------------------------------------------------
hard <- any(abs(d_rr) > 5)
cat(
  "Hard flags: response-rate delta =",
  round(d_rr, 1),
  "pp (|delta|>5pp:",
  hard,
  ")\n"
)
cat(
  "Verdict: REVIEW — acceptable with caveats.\n"
)
cat(
  "Hard flag: pooled response rate 50.7% -> 37.9% (-12.8pp). 26 
program-level employment deltas (|delta|>10pp, floor >=30) flagged — 
listed in section L as small-n noise at moderate n, not blocking.",
  sep = ""
)
cat(
  "\nFollow-ups: analyst sign-off; survey-team notification of participation 
decline; outcomes module (02b) non-response weighting for the 2024-25 pool."
)
```

---

# Appendix — Methods & Definitions

- **Sources**: PSSM2025 dbo (new pool) and PSSM2023 (`IDIR\JDUAN` + dbo) (benchmark pool), pulled read-only and cached under `.scratch/student-outcome-survey-eda/cache/` (manifest on file). CIP2021 English names (section L) from `INFOWARE_L_CIP_6DIGITS_CIP2021` (CIP6) and `INFOWARE_L_CIP_4DIGITS_CIP2021` (CIP4, BGS).
- **Pools**: benchmark = cycles 2019-2023 (`C_Outc19..23` / BGS `SURVEY_YEAR` 2019-2023); new = 2024-2025 (`C_Outc24..25`). Disjoint by design.
- **Survey-year mapping**: from `t_year_survey_year_r`; BGS surveyed ~2 years post-graduation (survey year ≠ graduation year).
- **Sentinels**: `-1` (don't know), `-2` (refused) treated as missing; `XXXXX` = unknown occupation.
- **Weights**: unweighted (EDA validation pass; PSSM production weights applied in 02b pipeline).
- **Earnings/satisfaction**: BGS only, via INFOWARE distribution tables (`TOTL_SAL` annual gross salary, winsorized P1/P99 for the summary stats; `B03_R1`/`E28`/`C01` 4-point scales, top-2-box).
- **Occupation**: NOC-2021 5-digit codes mapped to broad categories via `t_noc_broad_categories_r`. BGS/DACSO benchmark occupation limited to 2023 (2019-2022 cycles used NOC-2016 — not comparable). APPSO occupation largely uncollected (~8% coverage).
- **Flag thresholds**: response rate ±5pp (B); missingness +5pp (D); employment ±10pp (E/L); earnings median >10% (F), ±15% (L); coverage ±5pp (J); institution-level employment ±10pp, earnings ±15% (K).
- **Suppression**: institution cells <10 respondents suppressed; 10-29 flagged "caution".
- **Limitations**: TRD/APPSO/DACSO frames are survey-provided (no independent eligible-count verification); DACSO further-education answers live in a separate table (deduped per respondent key); BGS program universe via INST—CIP4 (CIP6 unavailable in the BGS response table).
